---
title: "Exploring VA Resumes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
```

```{r load data}
cert <- read.csv("~/git/stem_edu/data/stem_edu/working/Virginia_Resumes/resume_certification_info_va.csv")
edu <- read.csv("~/git/stem_edu/data/stem_edu/working/Virginia_Resumes/resume_education_info_va.csv")
job <- read.csv("~/git/stem_edu/data/stem_edu/working/Virginia_Resumes/resume_job_info_va.csv")
main <- read.csv("~/git/stem_edu/data/stem_edu/working/Virginia_Resumes/resume_personal_info_va.csv")
skill <- read.csv("~/git/stem_edu/data/stem_edu/working/Virginia_Resumes/resume_skill_info_va.csv")
```

for my own use: 

```{r size, echo=FALSE}
tibble(
  "Set" = c("Personal", "Jobs", "Education", "Certifications", "Skills"),
  "# Obs" = prettyNum(c(nrow(main), nrow(job), nrow(edu), nrow(cert), nrow(skill)), big.mark=",",trim = TRUE),
  "Ratio" = round((c(nrow(main), nrow(job), nrow(edu), nrow(cert), nrow(skill)))/nrow(main), 1),
  "# Vars" = c(ncol(main), ncol(job), ncol(edu), ncol(cert), ncol(skill))
)

tibble(colnames(job), class(colnames(job)))

job$jobISOStartDate[1]
dates <- lubridate::ymd(job$jobISOStartDate)
dates <- dates[!is.na(dates)] 
min(dates) #1957
max(dates) #2018

```

```{r scratch paper}
head(skill)
```


#### Data Review
##### BGT Job-Ad Data

Burning Glass Technology offers six principal domains within its Virginia resume data across the years 1957-2018.

* Personal - the base table of the resume, contains 10 columns of candidate descriptors including location and gender.
* Jobs - a related table of a candidate's job history including ONET code, job location, company sector, and dates.
* Education - a related table indicating level of study and, where possible, the school and field of study details.
* Certifications - a related table of certifications amd licenses associated with candidate.
* Skill - a related table indicating skills associated with job postings (3 fields for skill names, 3 for skill descriptors)
    Domain     |   Type  | Columns  | Rows (M) | Notes
---------------|---------|----------|----------| ----
Main           | base    | 10       | 0.5      | 
Jobs           | related | 10       | 3.9      |
Education      | related | 13       | 0.9      |
Certifications | related | 5        | 0.5      |
Skill          | related | 9        | 16.4     |






