---
title: "Main profile"
author: "Eirik Iversen"
date: "4/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(DataExplorer)
library(dplyr)
library(magrittr)
library(data.table)
library(ggplot2)
library(sdalr)
library(DBI)
```


```{r load data}
#load the RDS in working - already 2017 july
main <- readRDS('../../../data/stem_edu/working/BGexplorevalidate/BGT_main_0717.RDS')

#get columns of interest
main <- select(main, "bgtjobid",'jobdate', 'occfam', 'occfamname', 'employer', 'city', 
               'state', 'county', 'fipsstate', 'fipscounty', 'fips', 'lat', 'lon')
main <- as.data.table(main)
```

```{r dataexplorer}
overall <- melt(DataExplorer::introduce(main))
colnames(overall) <- c("Metric", "Value")
knitr::kable(overall)
missing <- DataExplorer::profile_missing(main)
colnames(missing) <- c("Column", "# Missing", "% Missing")
knitr::kable(missing)
```

```{r uniqueID}
### Uniqueness of Unique ID

#IDs that appear more than once
dupejobids <- main %>% dplyr::group_by(bgtjobid) %>% dplyr::summarise(count = dplyr::n()) %>% dplyr::filter(count > 1)

#Multi-appearance ID by date
dupecount <- main[bgtjobid %in% dupejobids$bgtjobid, c("bgtjobid", "jobdate")] %>% group_by(bgtjobid, jobdate) %>% summarise(count = n())

nonuniqueIDs <- nrow(dupejobids) - nrow(dupecount)
cat(paste(nonuniqueIDs, "unique job ids have multiple dates associated with them.") )
```

## HISTOGRAMS

Job Date
```{r jobdate hist}
##Job Date
values <-  main[!is.na(lubridate::as_date(jobdate)), lubridate::as_date(jobdate)]
unique_values <- length(unique(values))
frequencies <- as.data.table(table( values))
ggplot(frequencies, aes(values, N)) +
  geom_bar(stat = "identity", fill ="paleturquoise4", width=.7) +
  ggtitle(paste("Job Date", "Value Distribution")) +
  xlab("Job Date") +
  ylab("N") +
  theme(panel.background=element_blank(), axis.text.x = element_text(angle = 90, hjust = 1))
```

Employer, occfam, and occfamname are text fields so let's look for empty and missing strings
```{r cip hist, echo = TRUE, inclue = TRUE}
### Certification 
emp <- main$employer
(sum(is.na(emp)) + sum(emp == 'na') +sum(emp == '')) / length(emp)


occ <- main$occfam
(sum(is.na(occ)) + sum(occ == 'na') +sum(occ == '')) / length(occ)


occname <- main$occfamname
(sum(is.na(occname)) + sum(occname == 'na') +sum(occname == '')) / length(occname)

```
Employer is around 21% missing (filled with 'na'), occfam and occfamname are around 3% missing


## Questions 

* If occfam is missing, is occfamname always missing as well?
* what naming standardizations have been done to the employer column? 
* How does burning glass handle subsidiaries?
